results

### 2020-10-25 analysis/008_subject_exploration.Rmd
update results

```{r}
jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))

hybrid_volume_subjects <- hybrid_volume %>%
  group_by(oa_sponsor_type, articles, issued_year, issn) %>% 
  summarise(n_oa = n_distinct(doi)) %>% 
  left_join(jn_subjects, by = "issn") %>%
  ungroup()

hybrid_volume_subjects %>%
 # filter(container_title == "Journal of the American College of Cardiology") %>%
  group_by(top_level, articles, issued_year) %>%
  summarise(n_oa = sum(n_oa)) %>%
  ungroup() %>%
  group_by(top_level) %>%
  summarise(n_oa = sum(n_oa),
            n = sum(articles), .groups = "drop") %>%
  mutate(prop = n_oa / n) %>%
  arrange(desc(prop))

@@

- high share of elseveir waived articles due to ome issue published in Nuclear and Particle Physics Proceedings: "37th International Conference on High Energy Physics (ICHEP)
(ICHEP 2014)2-9 July 2014, Valencia, Spain" Do we want to exclude this journal?

@@

```{r}
field_funder <- hybrid_df %>%
  filter(oa_sponsor_type == "FundingBody") %>% 
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
   mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_reorder(oa_sponsor, n_oa_type))
ggplot(field_funder, aes(x = oa_sponsor, 
                         y = reorder(top_level, desc(top_level)), 
                         colour = prop, size = n_oa_type)) + 
  geom_point() +
  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
                      option = "magma",
                      name = "Percentage per funder",
                      breaks =  scales::extended_breaks()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid(font_size = 10) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank()) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", legend.box = "vertical") 
```

>>>>

```{r}
# jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))
# 
# hybrid_volume_subjects <- hybrid_volume %>%
#   group_by(oa_sponsor_type, articles, issued_year, issn) %>% 
#   summarise(n_oa = n_distinct(doi)) %>% 
#   left_join(jn_subjects, by = "issn") %>%
#   ungroup()
# 
# hybrid_volume_subjects %>%
#  # filter(container_title == "Journal of the American College of Cardiology") %>%
#   group_by(top_level, articles, issued_year) %>%
#   summarise(n_oa = sum(n_oa)) %>%
#   ungroup() %>%
#   group_by(top_level) %>%
#   summarise(n_oa = sum(n_oa),
#             n = sum(articles), .groups = "drop") %>%
#   mutate(prop = n_oa / n) %>%
#   arrange(desc(prop))
#

@@

- high share of elseveir waived articles due to ome issue published in Nuclear and Particle Physics Proceedings: "37th International Conference on High Energy Physics (ICHEP)
(ICHEP 2014) 2-9 July 2014, Valencia, Spain" Do we want to exclude this journal?

@@

```{r}
#short funder labels
funder_codes <- tribble(
  ~oa_sponsor, ~oa_sponsor_short,
  "Engineering and Physical Sciences Research Council", "EPSRC",
  "VSNU", "VSNU",
  "Wellcome Trust", "Wellcome",
  "European Research Council", "ERC",
  "Medical Research Council", "MRC",
  "Natural Environment Research Council", "NERC",
  "Biotechnology and Biological Sciences Research Council", "BBSRC",
  "Bill & Melinda Gates Foundation", "Gates",
  "Economic and Social Research Council", "ESRC",
  "Norway Institutes", "NO",
  "Other", "Other")
```

```{r}
field_funder <- hybrid_df %>%
  filter(oa_sponsor_type == "FundingBody") %>% 
  left_join(funder_codes, by = c("oa_sponsor_name"="oa_sponsor")) %>% 
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
  mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
ggplot(field_funder, aes(x = factor(oa_sponsor), 
                         y = reorder(top_level, desc(top_level)), 
                         colour = prop, size = n_oa_type)) + 
  geom_point() +
  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
                      option = "magma",
                      name = "Percentage per funder",
                      breaks =  scales::extended_breaks()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid(font_size = 10) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank()) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", legend.box = "vertical") 
```

### 2020-10-25 analysis/results.Rmd
update results

On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model. Of those, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_oa_volume %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio. These journals published a total of `r sum(hybrid_oa_volume$oa)` immediate open access articles between 2015 and 2019, corresponding to a share of `r round(sum(hybrid_oa_volume$oa) / sum(hybrid_oa_volume$articles) * 100, 1)`%. Table present the high-level findings by year, illustrating a moderate growth in hybrid open access. Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r hybrid_oa_volume %>% filter(issued_year == "2015") %>% pull(oa) %>% sum()` in 2015 to `r hybrid_oa_volume %>% filter(issued_year == "2019") %>% pull(oa) %>% sum()` in 2019. However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r round(ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake) * 100, 1)`% to `r round(ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake) *100, 1)`%, because the number of articles published also increased during this five-years period.

@@

>>>>

On 23 May 2020, Elsevier APC pricing list comprised `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` journals that supported the hybrid model, without reverse-flips: `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() - filter(flipped_jns_ind, oa_model != "Open Access")$n`.  Of those, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` hybrid journals published at least one open access article under a Creative Commons license between 2015 and 2019, corresponding to about `r round(hybrid_oa_volume %>% distinct(issn) %>% nrow() / els_jns_df %>% filter(oa_type == "Hybrid") %>% nrow() * 100, 0)`% of journal titles in Elsevier’s hybrid journal portfolio. These journals published a total of `r sum(hybrid_oa_volume$oa)` immediate open access articles between 2015 and 2019, corresponding to a share of `r round(sum(hybrid_oa_volume$oa) / sum(hybrid_oa_volume$articles) * 100, 1)`%. Table present the high-level findings by year, illustrating a moderate growth in hybrid open access. Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r hybrid_oa_volume %>% filter(issued_year == "2015") %>% pull(oa) %>% sum()` in 2015 to `r hybrid_oa_volume %>% filter(issued_year == "2019") %>% pull(oa) %>% sum()` in 2019. However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r round(ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake) * 100, 1)`% to `r round(ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake) *100, 1)`%, because the number of articles published also increased during this five-years period.

@@

### Field cultures

```{r}
hybrid_subject <- readr::read_csv(here::here("data", "jn_scopus_ind_subjects.csv"))
```

```{r}
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")

median_all <- subject_all_years %>% 
  ungroup() %>% 
  distinct(issn, .keep_all = TRUE) %>% 
  mutate(prop = n_oa /n) %>% 
  .$prop %>% 
  median()
# rank
oa_per_subject_rank <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  mutate(prop = n_oa / n) %>%
  ungroup() %>% 
  group_by(description) %>%
  summarise(median_prop = median(prop)) %>%
  arrange(desc(median_prop))

ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05))
  ) +
  coord_flip(ylim=c(0,0.2)) +
  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
  theme_minimal_vgrid(font_size = 10) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 2))
```

We used the Scopus’ All Sciences Journal Classification (ASJC) to examine if and to what extent open access uptake rates varied by subject. Figure presents the distribution of the journal’s five-years open access uptake rates by subject fields in comparison to the overall average (Mdn = `r round(median_all * 100, 1)`. The box plot reveals large variations between 27 journal subject fields (excluding one journal belonging to the area Multidisciplinary), with the highest OA rates in Immunology and Microbiology (Mdn = `r round(filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop) * 100,1)`), Environmental Science (Mdn = `r round(filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop) * 100,1)`) and Neuroscience  (Mdn = `r round(filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop) * 100,1)`), Environmental Science (Mdn = `r round(filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop) * 100,1)`) , and the lowest in Chemistry (Mdn = `r round(filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop) * 100,1)`), Materials Science (Mdn = `r round(filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop) * 100,1)`) and Dentistry (Mdn = `r round(filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop) * 100,1)`). Figure also highlights variations across subject areas. Notably, fields belonging to the subject areas Health Sciences and Physical Sciences show a lower median proportion of open access articles, whereas most Life Sciences and Social Sciences fields ranked above average.

```{r}
hybrid_jn_volume <- hybrid_oa_volume %>%
  group_by(issn) %>% 
  summarise(n_oa = sum(oa))

jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))

hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
funder_shares <- hybrid_df %>%
   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_jn_volume, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup() %>%
  group_by(top_level, oa_sponsor_type) %>%
  summarise(oa_type = sum(n_oa_type)) %>%
  mutate(prop = oa_type / sum(oa_type))
# overall props
oa_sponsor_type_all <- hybrid_df %>% 
  #exclude multidisciplinary
  filter(issn != "2095-9273") %>%
  count(oa_sponsor_type) %>% 
  mutate(prop = n / sum(n))

funder_shares %>% 
  filter(oa_sponsor_type %in% c("Author", "Agreement"), top_level != "Multidisciplinary") %>%
  pivot_wider(id_cols = top_level, names_from = oa_sponsor_type, values_from = prop) %>%
  filter(!is.na(top_level)) %>%
  ggplot() +
   ggalt::geom_dumbbell(
    aes(y = reorder(top_level, Author), x = Agreement, xend = Author),
    colour = "grey20",
    colour_xend = "#EDAE49",
    colour_x = "#30638E",
    size_x = 1.5,
    alpha = 0.9,
    size_xend = 1.5,
    show.legend = TRUE
   ) +
  geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "Author") %>% .$prop,  colour = "#EDAE49", linetype ="dashed", size = 1) +
    geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "FundingBody") %>% .$prop,  colour = "#30638E", linetype ="dashed", size = 1) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0,1)
  ) +
  labs(y = NULL, x = "Percentage Invoicing Type") +
  theme_minimal_vgrid(font_size = 10)
```

```{r}
subject_all <- funder_shares %>%
  group_by(top_level) %>%
  summarise(oa = sum(oa_type))
funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level")
# gt
funder_overview_subject %>%
  pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
  ungroup() %>% 
  # remove na and multidisciplinary 
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  relocate(Agreement, .after = Author) %>%
  arrange(desc(Author)) %>%
  gt::gt(rowname_col = "top_level") %>%
  gt::cols_label(oa = "OA Articles") %>%
  fmt_percent(
    columns = vars(Author, Agreement, `Fee Waived`, Other),
    decimals = 0) %>%
  cols_width(
    2:6 ~ px(80)) %>%
  cols_align(
    align = "right",
    columns = 2:6
  ) %>%
  cols_width(
    2 ~ px(120)
  ) %>%
    gt::cols_align(
    align = "left",
    columns = 1
  ) %>%
  fmt_number(
    columns = vars(oa),
    decimals = 0
  ) %>%
  data_color(
    columns =3:6,
    colors = scales::col_numeric(
      # custom defined values - notice that order matters!
      palette = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
      domain = NULL
    )
  ) %>%
  tab_spanner(
    label = md("**OA Sponsor Type**"),
    columns = 3:6
  ) %>%
    tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_column_labels(everything())
    )) %>%
   tab_options(
    column_labels.border.top.color = "white",
    column_labels.border.top.width = px(3),
    column_labels.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.bottom.color = "white",
    table.border.bottom.width = px(3)
      )
```

funder by subject

```{r}
#short funder labels
funder_codes <- tribble(
  ~oa_sponsor, ~oa_sponsor_short,
  "Engineering and Physical Sciences Research Council", "EPSRC",
  "VSNU", "VSNU",
  "Wellcome Trust", "Wellcome",
  "European Research Council", "ERC",
  "Medical Research Council", "MRC",
  "Natural Environment Research Council", "NERC",
  "Biotechnology and Biological Sciences Research Council", "BBSRC",
  "Bill & Melinda Gates Foundation", "Gates",
  "Economic and Social Research Council", "ESRC",
  "Norway Institutes", "NO",
  "Other", "Other")
```

```{r}
field_funder <- hybrid_df %>%
  filter(oa_sponsor_type == "FundingBody") %>% 
  left_join(funder_codes, by = c("oa_sponsor_name"="oa_sponsor")) %>% 
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_short), "Other", oa_sponsor_short)) %>%
  mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_relevel(oa_sponsor, "Other", after = Inf))
ggplot(field_funder, aes(x = factor(oa_sponsor), 
                         y = reorder(top_level, desc(top_level)), 
                         colour = prop, size = n_oa_type)) + 
  geom_point() +
  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
                      option = "magma",
                      name = "Percentage per funder",
                      breaks =  scales::extended_breaks()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid(font_size = 10) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank()) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", legend.box = "vertical") 
```
